---
title: netvis
output:
  github_document:
    number_sections: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

<!-- badges: start -->
<!-- badges: end -->

The goal of this repo is to demonstrate different visualisation techniques for communicating information about transport networks.

```{r, eval = FALSE, echo=FALSE}
# install.packages("devtools")
devtools::install_github("ITSLeeds/netvis")
```

# Example data

We'll use a route network dataset from the Propensity to Cycle Tool (PCT) to demonstrate the package. The PCT is a web application that allows users to explore cycling potential across England and Wales. The PCT is available at [www.pct.bike](https://www.pct.bike/).

# Static data visualisation

A simple visualisation of the data in a multi-panel static map is shown below.

```{r, echo=FALSE}
rnet_leeds = pct::get_pct_rnet("west-yorkshire")
rnet_leeds = rnet_leeds |>
  select(-1)
zones_leeds = zonebuilder::zb_zone("Leeds")
central_leeds = zones_leeds |>
  dplyr::filter(circle_id == 1) |>
  sf::st_union()
rnet_central = sf::st_intersection(rnet_leeds, central_leeds)
plot(rnet_central)
# sf::st_write(rnet_central, "test-data/rnet_central.geojson", delete_dsn = TRUE)
```

# Interactive data visualisation with line widths

The `tmap` package provides a simple way to create interactive maps. The code below shows how to create an interactive map of the route network data.

```{r, echo=FALSE}
m1 =
  tm_shape(rnet_central) +
  tm_lines(
    lwd = "bicycle",
    scale = 9
  )
tmap_save(m1, "maps/m1.html")
# browseURL("maps/m1.html")
webshot2::webshot("maps/m1.html")
width_multiplier = round(max(rnet_central$dutch_slc) / max(rnet_central$bicycle))
```

The maximum level of flow in the Go Dutch scenario is `r round(width_multiplier)` times the maximum level of flow in the baseline scenario. We can use this to scale the line widths as illustrated below.

```{r, echo=FALSE, out.width="50%", fig.show="hold"}
m2 = tm_shape(rnet_central) +
  tm_lines(
    lwd = "dutch_slc",
    scale = 9 * width_multiplier
  )
tmap_save(m2, "maps/m2.html")
# browseURL("maps/m2.html")
webshot2::webshot("maps/m1.html")
webshot2::webshot("maps/m2.html")
```

There are two problems with line widths in the maps shown above:

- The thinnes lines are too thin
- The thickest lines are too thick

```{r, echo=FALSE}
maximum_width = 15
minimum_width_multiplier = 5
```

Given that the maximum width is determined by the `scale` argument, we can solve the first problem by increasing the value of `scale`.
The second problem can be solved with a multiplier that prevents lines being x times thicker than the thinnest lines, `r minimum_width_multiplier` times in this case.

```{r, echo=FALSE}
# get 95th percentile of line widths
combined_values = rnet_central |>
  sf::st_drop_geometry() |>
  sapply(function(x) {
    quantile(x, 0.95)
  }
  )
minimum_value_allowed = max(combined_values) / minimum_width_multiplier
```

```{r, echo=FALSE, out.width="50%", fig.show="hold"}
scale_bicycle = maximum_width / (quantile(rnet_central$dutch_slc, 0.95) / quantile(rnet_central$bicycle, 0.95))
# summary(rnet_central$bicycle)
rnet_bicycle =   rnet_central |>
    mutate(lwd = case_when(
      bicycle < minimum_value_allowed ~ minimum_value_allowed,
      TRUE ~ bicycle
    ))
# summary(rnet_bicycle$lwd)
m3 = tm_shape(
  rnet_bicycle
 ) +
  tm_lines(
    lwd = "lwd",
    scale = scale_bicycle
  )
tmap_save(m3, "maps/m3.html")
# browseURL("maps/m3.html")
webshot2::webshot("maps/m3.html")

# For the Go Dutch scenario
scale_dutch = maximum_width
m4 = tm_shape(
  rnet_central |>
    mutate(lwd = case_when(
      dutch_slc < minimum_value_allowed ~ minimum_value_allowed,
      TRUE ~ dutch_slc
    ))
 ) +
  tm_lines(
    lwd = "lwd",
    scale = scale_dutch
  )
tmap_save(m4, "maps/m4.html")
# browseURL("maps/m4.html")
webshot2::webshot("maps/m4.html")
```

A function that does this work would have the following structure in Python:

```{python}
def scale_line_widths(
    data,
    baseline_column,
    dutch_column,
    maximum_width = 15,
    minimum_width_multiplier = 5
):
    """
    Scale line widths for interactive maps.

    Parameters
    ----------
    data : pandas.DataFrame
        Dataframe containing the data to be plotted.
    baseline_column : str
        Name of the column containing the baseline values.
    dutch_column : str
        Name of the column containing the Go Dutch values.
    maximum_width : int, optional
        Maximum width of the lines in the interactive map. The default is 15.
    minimum_width_multiplier : int, optional
        The minimum width of the lines is the maximum value in the baseline
        data divided by this value. The default is 5.

    Returns
    -------
    data : pandas.DataFrame
        Dataframe with a new column containing the scaled line widths.

    """
    # get 95th percentile of line widths
    combined_values = data[[baseline_column, dutch_column]].quantile(0.95)
    # Set mininum value to the maximum value in the combined data divided by the minimum width multiplier
    minumum_value_allowed = combined_values.max() / minimum_width_multiplier
    # scale line widths
    scale = maximum_width / (combined_values[dutch_column] / combined_values[baseline_column])
    data["lwd_dutch"] = data[dutch_column].apply(lambda x: minimum_value_allowed if x < minimum_value_allowed else x)
    data["lwd_bicycle"] = data[baseline_column].apply(lambda x: minimum_value_allowed if x < minimum_value_allowed else x)
    return data, scale
```

Starting with the data in `test-data/rnet_central.geojson` let's  test this function on a small subset of the data.

```{python}
import pandas as pd
import geopandas as gpd
import numpy as np
data = gpd.read_file("test-data/rnet_central.geojson")
data = data[["bicycle", "dutch_slc"]]
data = data.iloc[0:10, :]
data, scale = scale_line_widths(data, "bicycle", "dutch_slc")
print(data)
```