---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
library(tidyverse)
```

# netvis

<!-- badges: start -->
<!-- badges: end -->

The goal of this repo is to demonstrate different visualisation techniques for communicating information about transport networks.

<!-- ## Installation

You can install the development version of netvis from [GitHub](https://github.com/) with:

-->

```{r, eval = FALSE, echo=FALSE}
# install.packages("devtools")
devtools::install_github("ITSLeeds/netvis")
``` 

## Example

We'll use a route network dataset from the Propensity to Cycle Tool (PCT) to demonstrate the package. The PCT is a web application that allows users to explore cycling potential across England and Wales. The PCT is available at [www.pct.bike](https://www.pct.bike/).

```{r, echo=FALSE}
rnet_leeds = pct::get_pct_rnet("west-yorkshire")
rnet_leeds = rnet_leeds |>
  select(-1)
zones_leeds = zonebuilder::zb_zone("Leeds")
central_leeds = zones_leeds |>
  dplyr::filter(circle_id == 1) |>
  sf::st_union()
rnet_central = sf::st_intersection(rnet_leeds, central_leeds)
plot(rnet_central)
sf::st_write(rnet_central, "test-data/rnet_central.geojson", delete_dsn = TRUE)
```

